Large-scale harmonic patterns clustering

### 2010-03-09 Paper/Paper.tex
recent improvs
result organization
NMF stuff started

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{data_sizes}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{fig:sizes}
\end{figure}

\iffalse
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
\fi
% RESULTS:    
% 1bar s4     avg.       mean dist
% 1K          2.30777    0.150370
% 5K          1.88822    0.145327
% 10K         2.00349    0.143567
% 50K         2.59714    0.142257
% 100K        1.89185    0.142195
% 250K        2.07822    0.141796
% 500K        1.88141    0.142081
% 1M          1.91990    0.142118
% 2M          2.01141    0.141596

% 1bar s4 first X samples
% 1K          2.05621    0.157142
% 5K          2.01956    0.150045
% 10K         1.95797    0.148357
% 50K         1.91257    0.146069
% 100K        1.98744    0.145765
% 250K        1.95362    0.145819
% 500K        2.00899    0.145772
% 1M          2.09856    0.145924
% 2M          1.91375    0.146065

% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
% 1K          2.10588    0.155474
% 5K          1.92358    0.149619
% 10K         2.00653    0.148056
% 50K         1.89443    0.146328
% 100K        1.92187    0.146004
% 250K        1.93067    0.146024
% 500K        1.91913    0.145962
% 1M          2.24550    0.145873
% 2M          2.01653    0.145955

% 2bar s8     
% 1K          3.27198    0.171300
% 5K          3.56894    0.165960
% 10K         2.61057    0.165063
% 50K         3.03945    0.162439
% 100K        2.89310    0.161588
% 250K        2.90660    0.161910
% 500K        2.69134    0.161233
% 1M          3.05344    0.161057
% 2M          2.74493    0.161068

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\# codes & \# samples & avg. dist \\ \hline \hline
1 & $500$ & \\
10 & $5$K &\\
50 & $25$K &\\
100 & $50$K &\\
200 & $100$K & \\
500 & $250$K & \\
1000 & $500$K & \\
5000 & $2.5$M & \\ \hline
\end{tabular}
\end{center}
\caption{{Code book size.}}
\label{tab:cbsize}
\end{table}

@@

>>>>

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.99\columnwidth]{data_sizes}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{fig:sizes}
\end{figure}

\iffalse
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
size & $5$K & $10$K & $50$K & $100$K & $500$K & $2$M  \\ \hline
distortion & $$ & $0.143$ & $0.142$ & $0.142$ & $0.142$ & $$\\ \hline
\end{tabular}
\end{center}
\caption{{Distortion for a code book of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a code or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $10$ thousands to $2$ millions. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.}}
\label{tab:datasize}
\end{table}
\fi
% RESULTS:    
% 1bar s4     avg_distOLD    avg_dist
% 1K          0.150370
% 5K          0.145327
% 10K         0.143567
% 50K         0.142257
% 100K        0.142195
% 250K        0.141796
% 500K        0.142081
% 1M          0.142118
% 2M          0.141596
% 1bar s4 first X samples
% 1K          0.157142
% 5K          0.150045
% 10K         0.148357
% 50K         0.146069
% 100K        0.145765
% 250K        0.145819
% 500K        0.145772
% 1M          0.145924
% 2M          0.146065
% 0bar s4 (and test set same, we could do 1bar s4 in test set...)
% 1K          0.155474
% 5K          0.149619
% 10K         0.148056
% 50K         0.146328
% 100K        0.146004
% 250K        0.146024
% 500K        0.145962
% 1M          0.145873
% 2M          0.145955
% 2bar s8     
% 1K          0.171300
% 5K          0.165960
% 10K         0.165063
% 50K         0.162439
% 100K        0.161588
% 250K        0.161910
% 500K        0.161233
% 1M          0.161057
% 2M          0.161068

\subsection{Codebook Size}
Results in Table \ref{tab:cbsize}.

\begin{table}
\begin{center}
\begin{tabular}{|l|l|c|}
\hline
\# codes & \# samples & avg. dist \\ \hline \hline
1 & $500$ & $0.195939$ \\
10 & $5$K & $0.161812$ \\
50 & $25$K & $0.147472$ \\
100 & $50$K & $0.142194$ \\
200 & $100$K & $0.137548$ \\
500 & $250$K & $0.131211$ \\
1000 & $500$K & \\ \hline
\end{tabular}
\end{center}
\caption{{Code book size.}}
\label{tab:cbsize}
\end{table}
% RESULTS:      avg_distsOLD    avg_dist
% code 1        0.195939
% code 10       0.161812
% code 50       0.147472
% code 100      0.142194
% code 200      0.137548
% code 500      0.131211

@@

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|}
\hline
\# codes & \# bars & avg. dist \\ \hline \hline
2 & $1$ & $0.058435$ \\
4 & $2$ & $0.054792$ \\
16 & $4$ & $0.055489$ \\
256 & $8$ & $0.054860$ \\ \hline
2 & $2$ beats & $0.046955$ \\
4 & $1$ & $0.$ \\
16 & $2$ & $0.$ \\
256 & $4$ & $0.$ \\ \hline
\end{tabular}
\end{center}
\caption{{Code book size in function of the number of bars to encode.
All experiments done wtih $256 \times 500 = 128$K samples randomly chosen.}}
\label{tab:size_pattern}
\end{table}
% RESULTS:           avg_distsOLD   avg_dist
%2 codes 1 bar       0.180750       0.058435
%4 codes 2 bars      0.183631       0.054792
%16 codes 4 bars     0.184078       0.055489
%256 codes 8 bars    0.183464       0.054860
%---------------
%2 codes 2 beats     0.166840       0.046955
%4 codes 1 bar       0.171495       0
%16 codes 2 bars     0.173767
%256 codes 4 bars    0.174169       0

@@

\section{Links with NMF}\label{sec:nmf}
In this section we discuss links with non-negative matrix factorization
\cite{Lee2000} (NMF), another method to build a codebook representing some data.
NMF finds two matrices $H$ and $W$ that approximate some data $D$ (one
sample per line) by
\[
D = W \times H + \epsilon
\]
where $\epsilon$ is some noise and $W$ and $H$ only have non-negative 
elements. If $W$ is $\#samples \times n$, $n$ is the codebook size.

Note that NMF does not solve the same problem. It tries to explain each pattern
as a combination of codes. Thus it does not perform clustering, though you
could cluster patterns using their encoding weights.

In all our experiments we use Matlab's implementation of NMF:
nmflib\footnote{implementation: \url{http://www.ee.columbia.edu/~grindlay/code.html}}.

\subsection{Training on our Codebook}
NMF provides us with a way of measuring how much information is retained
in our code book. More specifically, we can train a code book of size $500$
using our online VQ algorithm, then decompose that code into $10$ basis
using NMF. We can compare that to $10$ basis from training NMF on the
whole data.

The idea is simple, if we encode our test set using basis from the
codebook as well as if we were using basis from the whole data, our codes
retain much of the information. That being said, the typical NMF implementations
are not expected to scale to data that does not fit in memory. Online VQ
could be an interesting preprocessing step. Future work include looking
at online NMF \cite{Thurau2009}.
